---
title: "ReadMe"
author: "Iván J. Barrera Olivera"
date: "27 de septiembre de 2015"
output: md_document
---
# Introduction
In this repo you could find my project of the Getting and Cleaning Data Course. In this project we were asked to create one R script called *run_analysis.R* that does the following:

1. Merges the training and the test sets to create one data set.
2. Extracts only the measurements on the mean and standard deviation for each measurement.
3. Uses descriptive activity names to name the activities in the data set
4. Appropriately labels the data set with descriptive variable names.
5. From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.

But, for my assignment, I decide do not follow the steps in the same order so I do the following:

1. Read each file of the test data and training data and assign the column names.
2. Merge the training and the test sets in one data frame.
3. Assign descriptive activity names to data set
4. Create new data set

# Read each file of the test data and training data and assign the column names

First, I download the file from the URL:

```{r, eval=FALSE,echo=TRUE, results='hide'}
# Download file
temp <- tempfile() # Create a temporal file
download.file ("https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip", temp)
unzip(temp) # Unzip temporal file
unlink(temp) # Delete temporal file
rm(temp) # Removes temporal file from Global Environment of R
```

Then, I read the *subject_test.txt* and *y_test.txt* files using `data.table`package, and assingning the corresponding column names for each one.

```{r, eval=FALSE, echo=TRUE, results='hide'}
# Reading and binding test data
library(data.table)
## Reading subject data and setting column name
subject.test <- fread("~/Documents/R Projects/Data_Science_Specialization/Cleaning data/Getting_Cleaning_Data_Project/UCI HAR Dataset/test/subject_test.txt")
colnames(subject.test) <- "subject"

## Reading activities data and setting column name
activities.test <- fread("~/Documents/R Projects/Data_Science_Specialization/Cleaning data/Getting_Cleaning_Data_Project/UCI HAR Dataset/test/y_test.txt", select = )
colnames(activities.test) <- "activities"
```


Then comes the **critical difference**, I create a vector with the number column with the mean and standard deviation for each measurement:

```{r, eval=FALSE, echo=TRUE, results='hide'}

## Creating a vector of columns with the mean and standard deviation for each measurement
variables <- c(1:6, 41:46, 81:86, 121:126, 161:166, 201:202, 214:215, 227:228, 240:241, 253:254, 266:271, 294:296, 345:350, 373:375, 424:429, 452:454, 503:504, 516:517, 526, 529:530, 542:543, 552, 555:561)
```

And I use this vector to read only the columns with measurements with mean and standard deviation of the *X_test.txt* file: 

```{r, eval=FALSE,echo=TRUE, results='hide'}

## Reading variable data
variable.test <- fread("~/Documents/R Projects/Data_Science_Specialization/Cleaning data/Getting_Cleaning_Data_Project/UCI HAR Dataset/test/X_test.txt", select = variables)
```

After this, I read the *feature.txt* that contain the variable names, filter it with the vector `variables` and use it to name the columns of the `variable.test`vector:

```{r, eval=FALSE,echo=TRUE, results='hide'}
### Renaming variable names
variable.names <- fread("~/Documents/R Projects/Data_Science_Specialization/Cleaning data/Getting_Cleaning_Data_Project/UCI HAR Dataset/features.txt")
variable.names <- variable.names$V2[variable.names$V1 %in% variables]
colnames(variable.test) <- variable.names

```


Finally, I bind my three data frames (`subject.test,activities.test,variable.test`) in one data frame (`test.data`).

```{r, eval=FALSE,echo=TRUE, results= 'hide'}
## Binding test data
test.data <- cbind(subject.test,activities.test,variable.test)
```

Then, I do the same with training data sets, as you can see in my *run_analysis.R* file.

So, as you can see, in this first step I extracts only the measurements on the mean and standard deviation for each measurement (original step 2) and label the data set with descriptive variable names (original step 4).

Also, I need to say that I use as "descriptive variable names" the same of the original data set, i.e. the labels of the features.txt file. I think this are the best labels because they were constructed using the main characteristics of each variable. For example, the majority of the variables start with a t (for time) or with a f (for frequency), then use Body or Gravity, then Acc(for acceleration) or Gyro (for gyroscope), then they specify if the variable refers to the jerk signals, the magnitude or both, then specify the type of statistical measure used to obtain it (e.g. mean, standard deviation, etc.) and, finally, identify the axis (XYZ)

# Merge the training and the test sets in one data frame
After read and assing column names for test and training data sets, I merge the two data frames (`test.data and train.data`) in one data frame (`smartlab.exp`), to do this I use `rbindlist()` of the `data.table` package, because is faster than `rbind()` of the base system.

```{r, eval= FALSE,echo=TRUE, results='hide'}
# Merging training and test sets
smartlab.exp <- rbindlist(list(train.data, test.data), use.names = TRUE)
```


# Assign descriptive activity names to data set
With my merged data frame (`smartlab.exp`), I assign the activity names using `mapvalues()`of the `plyr` package.

```{r, eval=FALSE,echo=TRUE, results='hide'}
# Assign descriptive activity names to data set
library(plyr)
smartlab.exp$activities <- mapvalues(smartlab.exp$activities,
                                     from = c("1", "2", "3", "4", "5", "6"),
                                     to = c("WALKING", "WALKING_UPSTAIRS", "WALKING_DOWNSTAIRS","SITTING","STANDING","LAYING"))
```

#Create new data set
Finally, I create my new and tidy data set using `aggegate()`, that enable us to group our data by subject and activities, and apply to the all variables the mean.

```{r, eval=FALSE, echo=TRUE}
# Creating new tidy data set
smartlab.exp.avg <- aggregate(smartlab.exp[,3:86, with = FALSE],
                              by = list(subject = smartlab.exp$subject,
                                        activities = smartlab.exp$activities),
                              FUN = mean)
```
